---
title: "Joint and Marginal Distributions"
author: "Le Wang"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Joint Distributions

Here we will play a dice game, **Craps**, as we did last semester, to illustrate concepts related to **Joint** and **Marginal** Distributions

```{r}
# Step 1: Simulate Data first 
set.seed(123456) 
x<-sample(1:6,100000, replace = T) 
y<-sample(1:6,100000, replace = T)
```

**Review Questions:**

1. Why `set.seed()`?
2. What does `sample(, , replace = T)` mean?
3. From the simulation design, can $x$ and $y$ be related?


```{r}
# Step 2: Calculate counts for each cell 
mytable<-table(x,y)
mytable
```


```{r}
# Step 2.b
# Alternatively, combine these two into a data frame 
data<-as.data.frame(cbind(x,y)) 
head(data)
```

```{r}
mytable<-xtabs(~ x+y, data=data)
mytable
```

From `R` documentation [here](https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/xtabs), note that `xtabs` is an `R` command to perform cross tabulation, which creates a contingency table (optionally a sparse matrix) from cross-classifying factors, usually contained in a data frame, using a formula interface with the cross-classifying variables (separated by `+`) on the right hand side (or an object which can be coerced to a formula). 

Interactions are not allowed. On the left hand side, one may optionally
give a vector or a matrix of counts; in the latter case, the columns are interpreted as corresponding to the levels of a variable. This is useful if the data have already been tabulated, as above. 


```{r}
# Step 3: Cell proportions 
prop.table(mytable)
```



# Visualization and Mosaic Plot 

We now use **Mosaic plot ** to visualize the joint distribution. To use the command for this purpose, we need to install the `vcd` package. 

```{r}
# Install the necessary packages first
install.packages("vcd")
```

## Equal Probability

Lets visualize the joint distribution with **equal probability** for each pair of $x$ and $y$ values.

```{r}
library(vcd)
mosaic(mytable, shade=TRUE, legend=TRUE)

```


## Unequal Probability

Lets simulate a case where the probabilities are unequally distributed over the space of the pairs of $x$ and $y$ values.


```{r}
set.seed(123456)
x<-sample(0:1,100000, replace = T, prob=c(.2,.8)) 
data<-data.frame(x, y=NA)
head(data)
```

Lets simulate some relationship between $x$ and $y$
```{r}
data$y[data$x == 1] <- sample(0:1, sum(data$x == 1), replace = T, prob=c(0.7,0.3)) 
data$y[data$x == 0] <- sample(0:1, sum(data$x == 0), replace = T, prob=c(0.1,0.9)) 
head(data)
```

Now, lets calculate the counts for each pair of values

```{r}
mytable<-xtabs(~ x+y, data=data)
mytable
```

Lets convert these counts to probabilities to obtain the joint distribution.

```{r}
prop.table(mytable)
```


Once we obtain the joint distribution (counts for each cell), we can now visualize it
```{r}
library(vcd)
mosaic(mytable, shade=TRUE, legend=TRUE)
```

What do we learn here? We learn two things (which combination is more likely):

1. $X = 1;Y = 0$ is most likely, while $X = 0;Y = 0$ least likely
2. We will use this to examine the conditional probability (defined later) as well.

# Joint Distribution to Marginal Distribution

```{r}
prop.table(mytable)
addmargins(prop.table(mytable))
```


