---
title: "A Detailed Example of Ethnicity Prediction"
author: "Le Wang"
output:
  beamer_presentation: default
  ioslides_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '/Users/lewang/Dropbox/02\ teaching/econ5043/02lectures/lecture2')
cnames<-read.csv("data/cnames.csv")
FLvoters<-read.csv("data/FLvoters.csv")
FLcensus<-read.csv("data/FLCensusVTD.csv")
dim(FLvoters)
FLvoters <- FLvoters[!is.na(match(FLvoters$surname,cnames$surname)),]
FLvoters<-na.omit(FLvoters)
cnames$pctothers<-100-(cnames$pctapi+cnames$pctblack+cnames$pctwhite+cnames$pcthispanic)
```


## Our problem

We would like to predict a voter's race. What is our procedure?

\pause
Step 1. Figure out the distribution

$$
\Pr[\text{Race}]
$$

\pause

$$
\begin{aligned}
& \Pr[\text{Race = white}], \Pr[\text{Race = black}], \Pr[\text{Race = hispanic}] \\
& \Pr[\text{Race = asian}], \Pr[\text{Race = others}]
\end{aligned}
$$

\pause 
Step 2. Pick the most likely value, in this case, always **white**


## More information

Now, suppose that I tell you this voter's last name is PIEDRA, what is your prediction for this person's race?

\pause
Step 1. Figure out the conditional distribution, i.e., the distribution for this subgroup

$$
\Pr[\text{Race} \mid \text{Surname}]
$$

\pause
In this specific case:
$$
\begin{aligned}
& \Pr[\text{race} = \text{white} & \mid \text{surname} = \text{PIEDRA}] \\
& \Pr[\text{race} = \text{black} & \mid \text{surname} = \text{PIEDRA}] \\
& \Pr[\text{race} = \text{hispanic} & \mid \text{surname} = \text{PIEDRA}] \\
& \Pr[\text{race} = \text{asian} & \mid \text{surname} = \text{PIEDRA}] \\
& \Pr[\text{race} = \text{others} & \mid \text{surname} = \text{PIEDRA}]
\end{aligned}
$$
\pause 
Step 2. Pick the most likely value and use it as our prediction. 


## How to do it?

```{r}
head(FLvoters)
```

---

```{r}
subset <- subset(FLvoters, surname == "PIEDRA")
prop.table(table(subset$race))
```

---

But in practice, if we do not have such information but have access to the census names data, what should we do?

```{r}
subset <- subset(cnames, surname == "PIEDRA")
subset
```

## How to evaluate our results?

For the white sample, we know that 

$$
\begin{aligned}
& \Pr[\text{race} = \text{white} & \mid \text{surname} = \text{PIEDRA}] \\
& \Pr[\text{race} = \text{black} & \mid \text{surname} = \text{PIEDRA}] \\
& \Pr[\text{race} = \text{hispanic} & \mid \text{surname} = \text{PIEDRA}] \\
& \Pr[\text{race} = \text{asian} & \mid \text{surname} = \text{PIEDRA}] \\
& \Pr[\text{race} = \text{others} & \mid \text{surname} = \text{PIEDRA}]
\end{aligned}
$$

$$
\max(\cdots) = \Pr[\text{race} = \text{white}  \mid \text{surname} = \text{PIEDRA}]
$$

---

```{r, size='tiny'}
subset
max(subset[, c("pctwhite","pctblack","pctapi","pctaian","pct2prace","pcthispanic")])
```

```{r, size='tiny'}
max(subset[, c("pctwhite","pctblack","pctapi","pctaian","pct2prace","pcthispanic")]) == subset$pctwhite
```


## Further Information

What if I tell you one more piece of information that this voter's last name is PIEDRA living in county 115 and VTD 66


\pause
Step 1. Figure out the conditional distribution, i.e., the distribution for this subgroup

$$
\Pr[\text{Race} \mid \text{Surname, County, VTD}]
$$

\pause
In this specific case:
$$
\begin{aligned}
& \Pr[\text{race} = \text{white} & \mid \text{surname} = \text{PIEDRA}, \text{county = 115, VTD = 66}] \\
& \Pr[\text{race} = \text{black} & \mid \text{surname} = \text{PIEDRA}, \text{county = 115, VTD = 66}] \\
& \Pr[\text{race} = \text{hispanic} & \mid \text{surname} = \text{PIEDRA}, \text{county = 115, VTD = 66}] \\
& \Pr[\text{race} = \text{asian} & \mid \text{surname} = \text{PIEDRA}, \text{county = 115, VTD = 66}] \\
& \Pr[\text{race} = \text{others} & \mid \text{surname} = \text{PIEDRA}, \text{county = 115, VTD = 66}]
\end{aligned}
$$
\pause 
Step 2. Pick the most likely value and use it as our prediction. 


---

```{r}
subset <- subset(FLvoters, surname == "PIEDRA" & county == 115 & VTD == 66)
prop.table(table(subset$race))
```

---

What if this dataset is not available? We have access to only census names data and Florida census data?

```{r}
head(cnames)
```

```{r}
head(FLcensus)
```

## Lets piece together information from two different datasets

$$
\begin{aligned}
& \Pr[\text{race}|\text{surname,residence}]\\
& =\frac{\Pr[\text{surname}|\text{race,residence}]\Pr[\text{race}|\text{residence}]}{\Pr[\text{surname}|\text{residence}]}
\end{aligned}
$$

\pause

$$
\begin{aligned}
&\Pr[\text{race} = \text{white}|\text{surname= PEIDRA,county=115, VTD=66}]\\
& =\frac{\overbrace{\Pr[\text{surname=PIEDRA}\mid \text{race=white,county=115, VTD=66}]}^{\textbf{Term A}} }{\underbrace{\Pr[\text{surname=PIEDRA}|\text{county=115, VTD=66}]}_{\textbf{Term C}}} \\
& \cdot \overbrace{\Pr[\text{race=white}|\text{county=115, VTD=66}]}^{\textbf{Term B}}
\end{aligned}
$$

## Term B

```{r}
subset <- subset(FLcensus, county == 115 & VTD == 66)
subset
```


```{r}
B <- subset[,"white"]
```

---

Let me also calculate this for other races

```{r}
B.black <- subset[,"black"]
B.hisp <- subset[,"hispanic"]
B.api <- subset[,"api"]
B.others <- subset[,"others"]
```

---

$$
\begin{aligned}
&\Pr[\text{race} = \text{white}|\text{surname= PEIDRA,county=115, VTD=66}]\\
& =\frac{\overbrace{\Pr[\text{surname=PIEDRA}\mid \text{race=white,county=115, VTD=66}]}^{\textbf{Term A}} }{\underbrace{\Pr[\text{surname=PIEDRA}|\text{county=115, VTD=66}]}_{\textbf{Term C}}} \\
& \cdot \overbrace{\Pr[\text{race=white}|\text{county=115, VTD=66}]}^{\textbf{Term B}}
\end{aligned}
$$

---

$$
\tiny
\begin{aligned}
& \Pr[\text{surname=PIEDRA}|\text{county=115, VTD=66}] \\
&= \Pr[\text{surname=PIEDRA}\mid \text{race=white,county=115, VTD=66}] \cdot \Pr[ \text{race=white}\mid \text{county=115, VTD=66}] \\
& + \Pr[\text{surname=PIEDRA}\mid \text{race=black,county=115, VTD=66}] \cdot \Pr[ \text{race=black}\mid \text{county=115, VTD=66}] \\
& + \Pr[\text{surname=PIEDRA}\mid \text{race=hispanic,county=115, VTD=66}] \cdot \Pr[ \text{race=hispanic}\mid \text{county=115, VTD=66}] \\
& + \Pr[\text{surname=PIEDRA}\mid \text{race=asisan,county=115, VTD=66}] \cdot \Pr[ \text{race=asian}\mid \text{county=115, VTD=66}] \\
& + \Pr[\text{surname=PIEDRA}\mid \text{race=others,county=115, VTD=66}] \cdot \Pr[ \text{race=others}\mid \text{county=115, VTD=66}] \\
\end{aligned}
$$

## Term A

This is my **Term A**:

$$
\Pr[\text{surname=PIEDRA}\mid \text{race=white,county=115, VTD=66}]
$$
\pause
I will make the following **conditional independence assumption**

$$
\begin{aligned}
& \Pr[\text{surname=PIEDRA}\mid \text{race=white,county=115, VTD=66}]\\
& = \Pr[\text{surname=PIEDRA}\mid \text{race=white}]
\end{aligned}
$$

---

Or similarly for every other racial category

$$
\begin{aligned}
& \Pr[\text{surname=PIEDRA}\mid \text{race=black,county=115, VTD=66}]\\
& = \Pr[\text{surname=PIEDRA}\mid \text{race=black}]
\end{aligned}
$$

---

In general 

$$
\Pr[\text{surname} \mid \text{race, residence}]=\Pr[\text{surname}\mid \text{race}]
$$


Does it make sense? Or, when will it be violated? 

\pause 

Conditional independence implies that once we know a voter's race, her residence location does not give us any additional information about her surname. 

\pause 
There is NO strong geographical concentration of certain surnames in Florida within a racial categroy. This will certainly be violated in the Chinese context. 

---

$$
\begin{aligned}
& \Pr[\text{surname=PIEDRA}\mid \text{race=white,county=115, VTD=66}]\\
& = \Pr[\text{surname=PIEDRA}\mid \text{race=white}] \\
\end{aligned}
$$

---

$$
\begin{aligned}
&\Pr[\text{race} = \text{white}|\text{surname= PEIDRA,county=115, VTD=66}]\\
& =\frac{\overbrace{\alt<2->{\Pr[\text{surname=PIEDRA}\mid \text{race=white}]}{\Pr[\text{surname=PIEDRA}\mid \text{race=white,county=115, VTD=66}]}}^{\textbf{Term A}} }{\underbrace{\Pr[\text{surname=PIEDRA}|\text{county=115, VTD=66}]}_{\textbf{Term C}}} \\
& \cdot \overbrace{\Pr[\text{race=white}|\text{county=115, VTD=66}]}^{\textbf{Term B}}
\end{aligned}
$$

---

Where do we have this information?
\pause

$$
\begin{aligned}
& \Pr[\text{surname=PIEDRA}\mid \text{race=white}]  \\
\pause 
& = \frac{\overbrace{\Pr[\text{race=white}\mid \text{surname = PIEDRA}]}^{\textbf{Term A.1}}}{\underbrace{\Pr[\text{race = white}]}_{\textbf{Term A.2}}} \cdot \overbrace{\Pr[\text{surname=PIEDRA}]}^{\textbf{Term A.3}}
\end{aligned}
$$

---

```{r}
A.1 <- cnames[which(cnames$surname=="PIEDRA"), "pctwhite"]/100
A.1
```

---

$$
\begin{aligned}
& \Pr[\text{surname=PIEDRA}\mid \text{race=white}]  \\
\pause 
& = \frac{\overbrace{\Pr[\text{race=white}\mid \text{surname = PIEDRA}]}^{\textbf{Term A.1}}}{\underbrace{\Pr[\text{race = white}]}_{\textbf{Term A.2}}} \cdot \overbrace{\Pr[\text{surname=PIEDRA}]}^{\textbf{Term A.3}}
\end{aligned}
$$

## Term A.2

```{r}
head(FLcensus)
```


```{r}
race.prop <- apply(FLcensus[,c("white", "black", "api", "hispanic", "others")], 
					2, 
					weighted.mean, 
					weights = FLCensus$total.pop)
A.2 <- race.prop["white"]
A.2
```

---

$$
\begin{aligned}
& \Pr[\text{surname=PIEDRA}\mid \text{race=white}]  \\
\pause 
& = \frac{\overbrace{\Pr[\text{race=white}\mid \text{surname = PIEDRA}]}^{\textbf{Term A.1}}}{\underbrace{\Pr[\text{race = white}]}_{\textbf{Term A.2}}} \cdot \overbrace{\Pr[\text{surname=PIEDRA}]}^{\textbf{Term A.3}}
\end{aligned}
$$

## Term A.3

```{r}
head(cnames)

```

---

```{r}
total.count<- sum(cnames$count)
A.3 <- cnames[which(cnames$surname == "PIEDRA"),"count"]/total.count
```



## Term A

```{r}
A <- A.1*A.3/A.2
A
```


---

Repeat this process for all other racial groups

```{r}
A.1.black <- cnames[which(cnames$surname=="PIEDRA"), "pctblack"]/100
A.1.hisp <- cnames[which(cnames$surname=="PIEDRA"), "pcthispanic"]/100
A.1.api <- cnames[which(cnames$surname=="PIEDRA"), "pctapi"]/100
A.1.others <- cnames[which(cnames$surname=="PIEDRA"), "pctothers"]/100

A.2.black <- race.prop["black"]
A.2.hisp <- race.prop["hispanic"]
A.2.api <- race.prop["api"]
A.2.others <- race.prop["others"]

A.black <- A.1.black*A.3/A.2.black
A.hisp <- A.1.hisp*A.3/A.2.hisp
A.api <- A.1.api*A.3/A.2.api
A.others <- A.1.others*A.3/A.2.others


A.black
A.hisp
A.api
A.others
```

---

$$
\begin{aligned}
&\Pr[\text{race} = \text{white}|\text{surname= PEIDRA,county=115, VTD=66}]\\
& =\frac{\overbrace{\Pr[\text{surname=PIEDRA}\mid \text{race=white,county=115, VTD=66}]}^{\textbf{Term A}} }{\underbrace{\Pr[\text{surname=PIEDRA}|\text{county=115, VTD=66}]}_{\textbf{Term C}}} \\
& \cdot \overbrace{\Pr[\text{race=white}|\text{county=115, VTD=66}]}^{\textbf{Term B}}
\end{aligned}
$$

## Term C

$$
\tiny
\begin{aligned}
& \Pr[\text{surname=PIEDRA}|\text{county=115, VTD=66}] \\
&= \Pr[\text{surname=PIEDRA}\mid \text{race=white,county=115, VTD=66}] \cdot \Pr[ \text{race=white}\mid \text{county=115, VTD=66}] \\
& + \Pr[\text{surname=PIEDRA}\mid \text{race=black,county=115, VTD=66}] \cdot \Pr[ \text{race=black}\mid \text{county=115, VTD=66}] \\
& + \Pr[\text{surname=PIEDRA}\mid \text{race=hispanic,county=115, VTD=66}] \cdot \Pr[ \text{race=hispanic}\mid \text{county=115, VTD=66}] \\
& + \Pr[\text{surname=PIEDRA}\mid \text{race=asisan,county=115, VTD=66}] \cdot \Pr[ \text{race=asian}\mid \text{county=115, VTD=66}] \\
& + \Pr[\text{surname=PIEDRA}\mid \text{race=others,county=115, VTD=66}] \cdot \Pr[ \text{race=others}\mid \text{county=115, VTD=66}] \\
\end{aligned}
$$

---

```{r}
residence <- subset(FLcensus, county== 115 & VTD ==66)
residence
```

---

```{r}
C <- A*residence["white"] + A.black*residence["black"] + A.hisp*residence["hispanic"]+ A.api*residence["api"]+A.others*residence["others"] 
C
```

---

```{r}
A*B/C
```



---

```{r}
cond.prob <- c(A*B/C, A.black*B.black/C, A.hisp*B.hisp/C, A.api*B.api/C, A.others*B.others/C)
names(cond.prob) <- c("white", "black", "hispanic", "api", "others")
cond.prob
```
