---
title: "Quiz 2"
author: "Your name here"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Write out the following covariance equation as a function of individual variances of $X$ and Yand their covariance. 

$$
Var(aX+bY+c)=
$$


2. True or False. No correlation implies that two variables are not depedent. 

**Answer:**

3. Conduct a test of correlation using the following data. Is there any statistically significant correlation between two variables?

```{r echo=FALSE}
set.seed(123456)
x <- rnorm(100)
y <- rnorm(100)
```

```{r}
# Conduct test here

```

**Answer:**


4. Multiple Choice (choose the best answer): If you have two stocks with identical risks (same variances) and invest the same amount of money in each stock, the risk is minimize when these two stocks are

(a) uncorrelated
(b) positively correlated
(c) negatively correlated
(d) all of above

**Answer:**

5. What is an unbiased forecast? 

**Answer:**

6. True or False. The combination of two unbiased forecasts is still an unbiased forecast.

**Answer:**

