---
title: "Quiz 3"
author: "Your name here"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
library(tidyverse)
```


Suppose that we have a hypothetical dataset as in the following table [imagine that we have already sorted it in some way]. Answer the following questions.

|   |   |   | 
|---|---|---| 
| Y | X | Z | 
|   |   |   | 
| 1 | 1 | 1 | 
| 1 | 1 | 1 | 
| 0 | 1 | 1 | 
| 0 | 1 | 0 | 
| 1 | 2 | 0 | 
| 0 | 2 | 0 | 
| 1 | 2 | 0 | 
| 0 | 2 | 1 | 
| 0 | 3 | 1 | 
| 0 | 3 | 1 | 
| 1 | 3 | 0 | 
| 0 | 3 | 0 | 

```{r}
data <- tribble(
  ~Y,~X,~Z,
  1,1,1,
  1,1,1,
  0,1,1,
  0,1,0,
  1,2,0,
  0,2,0,
  1,2,0,
  0,2,1,
  0,3,1,
  0,3,1,
  1,3,0,
  0,3,0,  
)
```


Lets follow the three topics that we learned in the past couple lectures. 

1. Figure out the **conditional distribution** of $Y$ given the predictors
2. Figure out what your predition of $Y$ will be given a value of your predictors based on a particular classifer, here **Bayes Classifier**.  
3. Why does your classifier would work? Figure out the mathematical justification behind this approach. 

## Conditional Distribution

1. What are the possible values for $Y$ here?

**Answer:** 

2. What is the conditional distribution of $Y$ given $X = 3$? In other words, what is the probability for every possible value of $Y$ for the subgroup with $X = 3$?

**Answer:** 

3. Lets verify your results not only for $X =3$, but for every subgroup using `prop.table()` with `margin` option.

```{r}
prop.table(table(data$X,data$Y), margin = 1)
```


## Bayes Classifier

4. Based on your results above, what is your prediction of $Y$ if $X = 3$?

**Answer:** 

5. Based on your results above, what is your prediction of $Y$ if $X = 1$?

**Answer:** 

## Why use Bayes Classifier?

6. For $X = 3$, if you use $Y = 0$ as your prediction, what is the chance that you will make an error?

**Answer:** 

7. For $X = 3$, if you use $Y = 1$ as your prediction, what is the chance that you will make an error?

**Answer:** 


